Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-12-15 paper/main.tex
Added discussion of possibly gain for Tahoe

If we look primarily at up- and download of the large files, we see that the candidates
that outperforms or are close to outperforming double \ac{SHA}-256, are BLAKE,
\ac{BMW}, Shabal, Simd and Luffa. The same is true if we eliminate the overhead
introduced by our bindings for the small files.

The fact that only these five candidates manages to outperform double
\ac{SHA}-256, is expected from the \ac{SUPERCOP} results, ref.
\autoref{app:supercop:bench}. However, the order and proportion of the
measurements are not. The reason for another order and relative speed is
primarily because we have not used the fastest implementation for all
candidates, and the relative speed of the chosen implementations compared to the
fastest implementation in \ac{SUPERCOP} differs, see Table
\ref{tbl:sha3:speedrelative}.

The performance of each candidate compared to the \ac{SHA}-256 version found in
Tahoe-\ac{LAFS} is also somewhat slower than you would expect. We blame this
mainly on our Cython-bindings and our skill. We believe a developer with more
knowledge of the inner workings of the implementations, the Python C-\ac{API}
and \ac{GCC} could have produced better results.

\subsection{Final round candidates}

BLAKE, Keccak, Skein, Grøstl and JH have made it through to the final round of
the \ac{SHA}-3 competition. From a performance perspective, only BLAKE actually
outperforms \ac{SHA}-256 in both the \ac{SUPERCOP} results and our measurements
for Tahoe-\ac{LAFS}. The other candidates are perhaps given an unfair treatment
in this paper, with the used implementations being relatively slow compared to
the best existing implementations. However, the reason for this is mentioned in
our criteria for selecting implementations.

None of the other four finalists outperform \ac{SHA}-256, and therefore it would be
more valuable for Tahoe-\ac{LAFS} to see the possible performance increase by
switching to a faster hash function implementation, i.e. BLAKE.

@@

@@

We have not used the fastest available implementations for all of the
candidates, we also know that our bindings creates an overhead for all calls.
Therefore we have tried to calculate the results that could have been achieved
in Tahoe-\ac{LAFS} if we had managed to obtain the speeds of the fastest
implementations in \ac{SUPERCOP}. The results can be found in
\autoref{appendix:optimized_sha3}.

% Describe inconsistencies in the numbers, and try to explain them.
% Time spent not hashing varies because: Variable use of CPU regs?
%
% Compare our results with general (SUPERCOP) results.
%
% How much gain can Tahoe-LAFS get by switching to SHA-3?

@@

\subsection{The use of Cython}

% TODO: Discuss the BLAKE run in Results when timing was done within Cython.

The generated C files from Cython used as Python bindings are not necessarily the most
optimized versions possible to write and they are ugly to read and change (from
a C perspective). However they are equal for all the \ac{SHA}-3 candidates
which means they should give the same performance penalty. The \ac{SHA}-256
version will of course not suffer from the same penalty. From
\autoref{tbl:cython:overhead} we can see the performance penalty of our
bindings, and we do believe better bindings can be made, also with the use of
Cython, but by someone with more skill and knowledge than us.

>>>>

If we look primarily at the large file up/download we see that the candidates
that outperforms, or are close to outperforming double \ac{SHA}-256 are BLAKE,
\ac{BMW}, Shabal, Simd and Luffa. The same is true if we eliminate the overhead
cause by our bindings for the small files. That only these five candidates
manages to outperform double \ac{SHA}-256 is expected from the \ac{SUPERCOP}
results (\autoref{app:supercop:bench}) however the order and proportion of the
measurements is not. This reason for another order and relative speed is
primarily because we have not used the fastest implementation for all
candidates, and the relative speed of the chosen implementation compared to the
fastest implementation in \ac{SUPERCOP} differs (Table
\ref{tbl:sha3:speedrelative}.  The performance of each candidate compared to
the \ac{SHA}-256 version found in Tahoe-\ac{LAFS} is also somewhat slower than
you would expect. We blame this mainly on our Cython-bindings and our skill. We
believe a developer with more knowledge of the inner workings of the
implementations, the Python C-\ac{API} and \ac{GCC} could have produced better
results.

\subsection{Final round candidates}
BLAKE, Keccak, Skein, Grøstl and JH have made it through to the final round of
the \ac{SHA}-3 competition. From a performance perspective only BLAKE actually
outperform \ac{SHA}-256 in both the \ac{SUPERCOP} results and our measurements
for Tahoe-\ac{LAFS}. The other candidates are perhaps given an unfair treatment
in this paper, with the implementations measures being relatively slow compared
to the best implementations. However the reason for this is mentioned in our
criteria for selecting implementations. Another obvious reason is that none of
the other four finalists outperform \ac{SHA}-256 and therefore it would be more
valuable for Tahoe-\ac{LAFS} to see the possible performance increase by
switching to a faster hash function implementation.

@@

\subsection{Possible gain for Tahoe-LAFS}

Even though not many of the
\ac{SHA}-3 implementation used by us managed to outperform the implementation
of double \ac{SHA}-256 in Tahoe-\ac{LAFS}, some did for large files, which
tells us there is a performance gain by switching which hash function is used
in Tahoe-LAFS. \autoref{tbl:tahoe:gain} compares the performance gains for
BLAKE, which did best in our measurements, versus \ac{SHA}-256, with overhead
caused by our measurements removed. We achieved as much as a 12,35\% gain when
uploading the 100MB files. We can see that the gain is smaller on downloads
than uploads, this is mostly caused by our Tahoe-\ac{LAFS} configuration, where
we need to upload shares to three servers, while we only need to download two
of them to get the original file.

\input{tables/tahoe_gain}

On another note it should be noted that most individuals will not have access
to the same high bandwidth network that our Tahoe-\ac{LAFS} servers ran on,
1Gbps or even 100Mbps. For an individual with a normal internet connection the
network is the bottleneck, and therefore the performance increase with a faster
hash algorithm implementation would not decrease the transfer time for large
files. If we consider Tahoe-\ac{LAFS} for organisations and enterprises which
typically empovers high bandwidth networks the change could indeed be
noticable.

\subsection{Faster \ac{SHA}-3 implementations will appear}

The obervation should also be made that \ac{SHA}-256 has probably been
optimized for the better part of a decade, while the \ac{SHA}-3 candidates are
relatively new. This will probably mean that in the future we will se even
faster versions of the \ac{SHA}-3 candidates, while it is more probable that
\ac{SHA}-2 is nearing the limit of its performance.

@@

We have not use the fastest available implementations for all of the
candidates, we also know that our bindings creates an overhead for all calls.
Therefore we have tried to calculate the results that could have been achieved
in Tahoe-\ac{LAFS} if we had managed to obtain the speeds of the fastest
implementations in \ac{SUPERCOP}. The results can be seen in
\autoref{appendix:optimized_sha3}.

% How much gain can Tahoe-LAFS get by switching to SHA-3?

@@

\subsection{The use of Cython}
The Python-bindings generated for Cython in C are not necessarily the most
optimized versions possible to write and they are ugly to read and change (from
a C perspective). However they are equal for all the \ac{SHA}-3 candidates
which means they should give the same performance penalty. The \ac{SHA}-256
version will of course not suffer from the same penalty. From
\autoref{tbl:cython:overhead} we can see the performance penalty of our
bindings, and we do believe better bindings can be made, also with the use of
Cython, but by someone with more skill and knowledge than us.
